---
title: "ex1: lm, var(beta)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# pb2: least sq regression

## a. run 100 times and save the slopes

```{r}
set.seed(21)
slopes <- numeric(100)
for(i in 1:100){
  x <- seq(1,40,1)
  y <- 2*x+1+5*rnorm(length(x))
  reg <- lm(y~x)
  slopes[i] <-coef(reg)[2]
}
```

## b. draw historgram of slopes


Theoritical variance: according to 1.4.2:

![](ex1/var(beta).png)

```{r}
par(mfrow=c(1,2)) # two plots in a row 
hist(slopes, freq=FALSE)

# And draw theorical distribution of slops:
X <- cbind(rep(1,40), x) # matrix X (intercept x0=1 is added)
XtX_inv <- solve(crossprod(X)) # theoritical s.d. of slopes
tsd = sqrt( 5^2 * XtX_inv[2,2] )
points <- seq(1.8,2.3,by=0.01)
lines(points, dnorm(points, mean=2, sd=tsd))
```

## c.  mean and empirical standard deviation of slopes
```{r}
mean(slopes)
sd(slopes)
summary(slopes)
```


## d. redo with with a skew, non-normal error distribution
```{r}
slopes2 <- numeric(100)
for(i in 1:100){
  x <- seq(1,40,1)
  y <- 2*x+1+5*(1-rchisq(length(x), df=1))/sqrt(2)
  reg <- lm(y~x)
  slopes2[i] <-coef(reg)[2]
}
hist(slopes2, freq = FALSE)
mean(slopes2)
sd(slopes2)
lines(points, dnorm(points, mean = 2, sd = tsd))
```

==> The approximation fit well because of the central limit theorem. 


# pb3 airline

## a. plot the data

```{r}
airline <- scan("http://stat.ethz.ch/Teaching/Datasets/airline.dat")
plot(airline, type="l")
```


## b. plot logarithms of the data
```{r}
logairline <- log(airline)
plot(logairline, type="l")
```

a better way to show time series data (x-axis times): 
```{r}
t <- ts(airline, start = c(1949,1), frequency = 12)
plot(c(time(t)), airline, xlab="Time", type = "l")
 ```


## c. define linear model
```{r}
f1 <- 1:144
f2 <- rep(c(1,rep(0,11)),12)
f3 <- rep(c(0,1,rep(0,10)),12)
f4 <- rep(c(0,0,1,rep(0,9)),12)
f5 <- rep(c(0,0,0,1,rep(0,8)),12)
f6 <- rep(c(0,0,0,0,1,rep(0,7)),12)
f7 <- rep(c(0,0,0,0,0,1,rep(0,6)),12)
f8 <- rep(c(rep(0,6),1,rep(0,5)),12)
f9 <- rep(c(rep(0,7),1,rep(0,4)),12)
f10 <- rep(c(rep(0,8),1,rep(0,3)),12)
f11 <- rep(c(rep(0,9),1,0,0),12)
f12 <- rep(c(rep(0,10),1,0),12)
f13 <- rep(c(rep(0,11),1),12)
```

a better way to define the fs: 
```{r}
trend <- 1:144
month <- as.factor(rep(month.name, 12))
reg2 <- lm(logairline ~ trend+month-1)
```

## d. fit linear model
```{r}
reg <- lm(logairline ~ f1+f2+f3+f4+f5+f6+f7+f8+f9+f10+f11+f12+f13-1)
par(mfrow=c(3,2))
trend <- 1:144
plot(trend, airline, type="l", main="original data")
plot(trend, logairline, type="l", main="log data")
plot(trend, fitted(reg), type="l", main="fitted data")
plot(trend, residuals(reg), type = "l", main="residuals")
plot(trend, rnorm(144), type="l", main="simualted normal data")
plot(trend, rnorm(144), type = "l", main="another simulated normal data")
```

