---
title: "ex2: lm(+assumption checking, model selection); NP density estimation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. SMSAs

## 1. inspect data
```{r}
url <- "http://stat.ethz.ch/Teaching/Datasets/mortality.csv"
mortality <- read.csv(url, header = TRUE)
mortality <- mortality[,-1]
```

use `splom()` in the lattice package to draw pairs plot: 

```{r}
# install.packages("lattice")
library(lattice)
splom(~mortality, pscales=0, cex=0.5)
```

The response variable mortality versus the predictor variables is the last row of the plot.

As the Pop, HC, NOx, SO2 have skewed distributions, we log-transform them. 

Now add log-transformation: 
```{r}
mortality_old <- mortality
mortality[,"logPop"] <- log(mortality[,"Pop"])
mortality[,"logHC"] <- log(mortality[,"HC"])
mortality[,"logNOx"] <- log(mortality[,"NOx"])
mortality[,"logSO2"] <- log(mortality[,"SO2"])
col_num <- which( names(mortality) %in% c("Pop", "HC", "NOx", "SO2") )
mortality <- mortality[,-col_num]
```


Now the scatter plot looks better: 
```{r}
splom(~mortality[,c("Mortality","logPop", "logHC", "logNOx", "logSO2")], pscales=0, cex=0.5, col="black")
```



Fit the data to linear model, fit the full model (i.e. use all predictor variables, starting point of backward selection): 
```{r}
mortal.full <- lm(Mortality~. , data=mortality)
```

Fit the empty model (starting point of forward selection):
```{r}
mortal.empty <- lm(Mortality~1, data=mortality)
```

###Tukey-Anscombe plot: fitted-value VS residuals. 

```{r}
plot(fitted(mortal.full), residuals(mortal.full),xlab="fitted value", ylab="residual", main="Tukey-Anscombe")
abline(h=0, lty=2)
```

==> linear assumption is valid. (error has 0 mean)

### QQ plot: empirical quantile VS standard quantiles
```{r}
qqnorm(resid(mortal.full), xlab="standard normal quantiles", ylab="empirical quantiles")
qqline(resid(mortal.full), lty=2)
```

==> the error is normal N(mu, sigma2), intercept=mu~0, slope=sigma

look at the fitting summary: 
```{r}
summary(mortal.full)
```
The summary statistic of the model reveals that only JanTemp, Tain, NonWhite, logNOx are significant on the response. 

## 2. stepwise variable selection

### backward selection
```{r}
mortal.bw <- step(mortal.full, dir="backward")
```

### forward selection
```{r}
motal.fw <- step(mortal.empty, dir="forward", scope = list(upper=mortal.full,lower=mortal.empty))
```

### all-subsets regression
Both fw and bw are greedy algorithms, in each step it adds variables and never deletes them, although **an addition of a new variable may render one or more of the already added variables non-significant.** 

Use the `leaps` package. 

```{r}
# install.packages("leaps")
library(leaps)
mortal.alls <- regsubsets(Mortality~., data=mortality, nvmax=9)

source("ftp://stat.ethz.ch/Teaching/maechler/CompStat/cp-plot.R")
p.regsubsets(mortal.alls, cex=0.8, cex.main=.8)
```



#2. mixture of two normal distributions
Generate the mixture-distributed data: 
```{r}
set.seed(1)
data <- numeric(100)
for(i in 1:100){
  p <- runif(1, min=0, max=1)
  if(p<0.2)
    data[i] <- rnorm(1, mean=0, sd=sqrt(0.01))
  else 
    data[i] <- rnorm(1, mean=2, sd=1)
}
hist(data)
```

another way to generate data withou loop: 
```{r}
n <- 100
data <- ifelse(runif(n, min=0, max=1)<0.2,rnorm(n, mean=0, sd=sqrt(0.01)), rnorm(n, mean=2, sd=1))
hist(data)
```


## a. best bandwidth with kernel density estimator

The desity estimator is in `density()`.

```{r}
ke <- density(data, bw=0.1, n=61, from=-1, to = 5)
summary(ke)
```

An estimated plug-in bandwidth is obtained by using `bw ="sj"`.


`ke$x` are the points at which the estimate is to be computed, we can use the average error to estimate the IMSE: 

```{r}
true_density <- 0.2*dnorm(ke$x, mean=0, sd=sqrt(0.01))+0.8*dnorm(ke$x, mean=2, sd=1)
quality <- mean((ke$y-true_density)^2)
```


## b. carry out simulation to evaluate quality of bandwidth

before doing the 200 simulations, create variables to hold the results: 

```{r}
n <- 100
bandwidths <- c(as.list(2^seq(-5,1,0.2)), "sj")
sj <- numeric(200)
qualities <- matrix(ncol=length(bandwidths), nrow=200)
```
For bandwidths we add the "sj" as last element. 

```{r}
for(i in 1:200){ # simulate 200 times
  data <- ifelse(runif(n, min=0, max=1)<0.2,rnorm(n, mean=0, sd=sqrt(0.01)), rnorm(n, mean=2, sd=1))
  for(j in seq_along(bandwidths)){
    ke <- density(data, bw=bandwidths[[j]], n=61, from=-1, to=5)
    qualities[i,j] <- mean((ke$y-true_density)^2)
    if(bandwidths[[j]]=="sj") {sj[i] <- ke$bw}
  }
}
```

find the results: 

```{r}
sj.mean <- mean(sj)
sj.mean

results <- apply(qualities, 2, mean) # take col mean
bandwidths[[which.min(results)]]
```

Then plot the bandwith VS quality curve, also plot the selected bandwidth selected with "sj" by a red point: 

```{r}
plot(2^seq(-5,1,0.2), results[-length(bandwidths)],xlab="bandwidth", ylab="quality", main="Guassian kernel result")
points(sj.mean, results[length(bandwidths)], col="red")
```

